Writing Trash Bin

### 2024-04-01 writingdumpster.qmd
sorted the misc notes in the writing dumpster

---
title: "Writing Trash Bin"
author: "Harriet Mason"
format: html
---
# Introduction
### Motivation for uncert vis
- @Suh2023 found that the main issue with deployment of AI was communication between data science and subject matter experts, since subject matter experts didn't want technical presentations that didn't explain exactly when and how a model would fail.
- (maybe include some of the noise vs signal stuff)

# Definition Notes
## General
- @uncertchap2022 People do not over react when an ensemble member barely misses their town and People overreact more when 1/9 enselbles appear to be hitting their town than when 1/33 ensembles.
- @uncertchap2022 Correll and Gleicher [9] found that violin plots (mapping probability to area or width at a given y position) and gradient plots (mapping probability to opacity) lead to more intuitive assessments of value likelihood and “ability to surprise,” which are more closely aligned with statistical defini- tions of uncertainty." wtf is the statistical definition of uncertainty
- @Thomson2005 titled their paper "A typology for visualizing uncertainty" however it is a taxonomy for the specific types of uncertainty that we may want to visualise and a way to calculate that uncertainty. While this paper is more about mathematics than visualisation, it is clearly written as a precursor to visualisation.
- @Griethe2006 commented that uncertainty appears at every step in the analyis pipeline, and commented on some the mathematical definitions that seek to quantify uncertianty, but did not explain how these mathematical definitions can be combined across levels. Unfortunately this quantification is needed since she mentioned that in order to visualise uncertainty, it needs to be measurable.
- mentions uncertainty can be variance, precision, accuracy, reliability, or related concepts @Hullman2016
- @uncertchap2022 "can enter the analysis pipeline during the mea- surement, modeling, and forecasting phases"

## Probability/variance/error etc (Cite Gap 3)
> Some consider it to be synonymous with specific terms defined in mathematics, such as probability [*HERE*], variance[*HERE*], error[*HERE*], or precision[*HERE*]. Others consider it to be an encompassing umbrella term of which for these mathematically defined objects are only examples[*HERE*], and include concepts that are somewhat related, such as missing values[*HERE*]

## Redefined uncertainty (cite gap 3)
> This wide array of uncertainty definitions becomes a snake eating its own tail, as many authors opt to redefine uncertainty themselves[*HERE*], ignore defining uncertainty entirely [*HERE*], or pick a definition from a seemingly randomly previously published papers [*HERE*], all of which make variance in uncertainty definitions even worse

#### Uncertainty Taxonomies (Cite Gap 7)
> Taxonomies split uncertainty based on an endless stream of arbitrary lines in the sand, such as whether the uncertainty is due to true randomness or a lack of knowledge [@Spiegelhalter2017 *HERE*], if the uncertainty is in the the attribute, spatial elements, or temporal element of the data [*HERE*], whether the uncertainty is scientific (e.g. error) or human (e.g. disagreement among parties) [*HERE*], if the uncertainty is random or systematic [*HERE*], statistical or bounded [], accuracy or precision [*HERE*], if the uncertainty is about the past or the future [*HERE*], the stage of the data analysis pipeline it comes from [*HERE*], the severity of the uncertainty [*HERE*], how quantifiable the uncertainty is [@Spiegelhalter2017], and many, many, many more examples.

- @uncertchap2022 mention aleatric, epistemic, ontological divide and also differentiates risk with uncertainty
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016

@@

- @Griethe2006 clearly sees uncertainty as additional information or data, and therefore sees the uncertainty visualisation problem as a specific case of the high dimensional visualisation issues.

# Noise vs Signal
- @uncertchap2022 "Effectively communicating uncertainty is necessary for establishing scientific transparency"
- @Griethe2006 defined uncertainty as "the degree to which the lack of knowledge about the amount of error is responsible for hesitancy in accepting results and observations without caution".

@@

### Value Extraction (Cite Gap 15)
- @uncertchap2022 Attribute substitution: people will substitute a hard mental calculation for an easy one (attribute substitution heuristic). Name "deterministic constructual error" when people try to substitute uncertainty information for deterministic information (e.g. error bars for high and low temperature even when the visualisation was explained - from icing study. This makes sense because temperature is GIVEN as high and lows so it is a natural interpretation). I think this result is in line with the idea that people give the best estimate based on what the visualisation suggests. i.e. people might have assumed you meant "what is the lowest possible temperature". There is a level of "asking a deterministic question for a probabilistic question". 
- @uncertchap2022 Quantile dot plot created so people dont have to visually calculate the area under the curve (so authors are aware of the mental challenge of some of these things) Decrease variance with respect to density plots (it makes sense because you can directly count and quantify raher than guesstimate) (both come from the same source, when-ish is my bus)
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016 
- @Correll2018 “place your 5 ships on the safest locations on the board.” (minimise danger) It is a convoluted value extraction task, where participants basicaly have to look at the palette and see that if it is in the lower uncertainty, and lower error it is better, but higher uncertainty is better
- @Ibrekk1987 expected participants to calculate the mean using a pie chart that had numerical bins combined (e.g. segment one was 0-2 inches of snow, and segment two was 2-4 inches of snow) so it could not be done by calculation. The authors also expected participants to use a CDF to calculate the mean, which needs to be converted *back* into a PDF to do so mathematically, so it was unclear how participants were supposed to do this calculation visually unless they were to just make a blind guess. The authors interpreted this as the participants misinterpreting how the CDF works, not seeming to realise they gave them an impossible task.
- @Sanyal2009 users performed significantly well using Glyph-size when the task was to search for locations of least uncertainty. However, both Glyph-color and Surface-color performed better than Glyph- size when the task was to search for locations of high uncertainty. We did not expect to find a significant difference between the two search tasks since both the tasks were designed to find extremes in the data
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014

### Decision making stuff (Cite Gap 16)
- @Boone2018 organised responses from participants by the approach they used
- @Hullman2016 suggested authors look for signs of heuristics in responses and ask subjects to describe their strategy in their response. 
- @Hullman2016 What was particularly interesting was people's tendancy to ignore uncertainty in a plot all together, and use a pattern in the signal (such as distance between means) to answer questions that require uncertainty information. 
- @Kim2019 believed an understanding of participants prior beliefs could assist in interpreting uncertainty visualisation results that tend to be incredibly noisy.
- @Daradkeh2015 found that people tended to to make decisions on a small number of salient pieces of information rather than on systematic evaluations of the costs and benefits of choices.
- @Daradkeh2015 found that people tended to ignore or give little weight probability information and instead focused on the severity of undersierable outcomes.
- believed that offering enough decisions with varying rewards you can observe the true subjective probability, however whether or not probability is the right representation of subjective uncertainty is a topic of debate @Hullman2016
- she also discusses graphical interface (users have to draw the probability distribution) and framing probabilities as natural frequencies are better choices because they reduce noise in responses [@Hullman2016]
- Consider incorporating utility functions. Include "probability-coherence" checks [@Hullman2016]
- evidence that people do not follow the axioms of probability when conveying subjective probabilities. How do you deal with this as an analyst (e.g. if the probabilities do not sum to 1 @Hullman2016
- people will discount information they perceive as discountable if they are able. Including mean estimates on depictions of mass can cause people to discount the uncertainty information and use the difference between means as a proxy for the probability distribution [@Kale2021].
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016

@@

### Trust stuff (Cite Gap 18)
- @uncertchap2022  mentions "other studies have found that people trust icon arrays more than other common visualization techniques" without commenting on whether or not that trust is good?

# Communication considerations
## Uncertainty
- @Spiegelhalter2017 it is better to understand your audience and iterate towards a final design, rather than only consider one option.
- Some markers of probability that have common uses, probabilities such as 0 (impossible), 0.5 (a fair coin flip), 1 (certain) are easy for people to have an internal sense of [@Hullman2016]
- It is incredibly hard to communicate small risks, because people cannot differentiate between 1/100 an 1/1000, and communicating low probability but high impact risks are particularly challenging [@Spiegelhalter2017]. This issue can be somewhat alleviated by communicating the likelihood of an event in relation to another event on a similar scale (such as being struck by lightning) [@Spiegelhalter2017]. 
- The framing of how we communicate risks matters due to the affect heuristic, so there will be a difference in peoples behaviours if, for example, you tell people they have a 90% chance of survival vs a 10% chance of death, and when expressing a frequency, a larger numerator communicates a larger risk [@Spiegelhalter2017]. 
- The reference class of a particular risk should be explicitly stated to avoid confusion, for example if we are communicating the chance of rain, the time period over which we are expecting this chance of rain (i.e. 50% chance today or this week) needs to be specified [@Spiegelhalter2017]. 
- industrial accidents have a "tollerable" risk chart based on risk and number of fatalities @Spiegelhalter2017
- People believe in the law of small numbers but dont believe in the lar of large numbers @Hullman2016
- experiencing an event makes your percieved probability higher than if you see a description of it @Hullman2016
- Frequency thing means that more effective communication to participants in a study means more effective science and we have a roll on effect @Hullman2016
- Probability judgements are less relative than other psychophysical judgements, 0%, 50% and 100% all have well understood meanings. Subjective probability distributions are spontaneous, and subjects are constructing a distribution upon being asked rather than articulating a fully formed mental representation of their beliefs. Therefore it is subject to heuristics. consider anchoring a scale with farmiliar probabilities . not sure if being asked for the entire distribution helps, e.g. the graphical one reduces noise but asking people for an interval estimation leads to overprecision and overconfidence @Hullman2016
- The big data paradox shrinks confidence intervals but magnifies bias [@Bradley2021] because the confidence intervals shrink due to the sample size but the imprecise target group creates bias.

## Visualisation
- @uncertchap2022 Icon array and quantile dot plots both support area and frequency based judgements.
- @uncertchap2022 frequency framing: (e.g. 1 out of 10 rather than 10%) better (plot: icon array, quantile dot plot). 
- @Spiegelhalter2017 even if most audience members can read a plot, they may not read it in the same way. People with high numeracy skills count pictographs, but people with low numeracy assess area, and seemingly unrelated elements may not equally influence all these methods of reading, for example scattering in a pictograph makes counting particularly more difficult. *also add that perceptual task audience thing*
- @Spiegelhalter2017 common sentiments such as avoiding chart junk, considering a table in place of a visualsiaiton, use of narrative labels, managing emotion, less is more approach, etc, are the main concepts expressed. 
- @Spiegelhalter2017 suggests that the graphic should allow "part-to-whole" comparison on an appropriate scale, which seems to be an extension of the framing concept discussed when speaking about uncertainty communication more broadly.
- @Hullman2016 discussed the role of heuristics in the analysis of results
- @Kinkeldey2014 stress the importance of selecting the correct audience, as numerical and visual literacy can differ vaslty between subpopulations so the results from one populations may not reflect the results from another. 
- we consider a summary table as a visualisation, illuminate graphics with words and numbers, design graphics to allow a part-to-whole comparison on an appropriate scale, use images and metaphors to gain/retain attention but don't arouse undue emotion, assume low numeracy of a general public audience and adopt a less-is-more approach by reducing the need for inference and making clear and explicit comparison, be cautious about interactivity and animations (may introduce unnecessary complexity), avoid chart junk (like 3d bar charts) @Spiegelhalter2017
- @Maceachren2012 found __????_ to be the worst aesthetics to map to uncertainty to as they don't have an intuitive interpretation.
- Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. 
- No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012]. 
- This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016].
- coloured maps for earthquake risk are easily interpretable @Spiegelhalter2017

# Main section outtakes
Conclusion
In this paper we have highlighted a series of misunderstandings, confusions, and methodical errors in the uncertainty literature. Many of these issues are not problems in isolation, so long as the authors themselves are aware of it and state these assumptions in their papers. Our problem with this body is work is that the authors themselves seem to be oblivious to this vast array of problems, considering not a single one of the problems discussed here are mentioned explicitly in a single visualisation paper.

It is clear that value retrival and questions with a specific ground truth are not appropriate, as uncertainty visualisation will simply follow the existing visualisation rules. Testing a secondary feature of a plot, that is, how strong a signal is becomes a little more complicated. Line up protocols have been used in adjacent work that look at the strength of the signal in a plot, and this idea of identifying if plots have some "barely noticiable differences" could be utilised.

Specifically, a definition that has a place for all the concepts that currently sit under the "uncertainty" umbrella, explains how these concepts relate to uncertianty, and also allows us to quantify them in a way where they can be combined or visualised. This work would not only benefit uncertainty visualisaiton, but also statistics more broadly.

The problem of trying to assess uncertainty without knowing *why* we are depiting uncertainty has been noticed by other authors. @Spiegelhalter2017 noted we "cannot assess the quality of risk communication unless the objectives are clear".

Does the flip (hypothesis test vs uncertainty visualisation) have mathematical implications in how the plots should be assessed? This implies a visible signal in an uncertainty plot should indicate some divergence from the null.

This increased sensisitivity could be the result of another aspect of uncertainty that is ignored. Statstical tests are typically built upon a series of assumptions, and it is difficult to identify *which* assumption failure caused the data to fail the statistical tests. Lineup plots allow us to see *how* our data is different to the assumed distribution, and better understand *why* our data may have failled a test ^[Did Patricks paper compare statistical tests to lineup protocols in the event an earlier assumption of the tests, such as, independent observation? I think he did check some. I wonder if you can quantify the information difference you get from a rejected statistical test vs a rejected line plot].

It is unclear how participants are supposed to incorperate uncertainty information into their responses when uncertainty visualisation authors themselves are unsure. Other authors interpret this difficulty as a combination of the uncertainty and task complexity [@Kinkeldey2014], and not as a general confusion. All these examples sever to illustrate that *we dont know how to measure uncertainty* and if we don't know how to measure it, *we cannot evaluate the quality of a visualisation*.

Using the most commonly presented methods In the absense of an encapsulating mathematical definition this will not change. That being said, even with a definition that allows us to combine these uncertainties into a single value, it would still be beneficial for authors to develop specialty visualisations for different sources of uncertainty, so analysts can utilise visual tools to understand the uncertainty in their data. still might be desirable to see which stage of the analysis is creating uncertainty in the visualisaiton.

@Hofmann2012 showed two distributions in 20 different visualisations (a lineup protocol) using a jittered sample, a density plot, a histogram, and a box plot and asked participants. Participants were asked to report in which of the plots was "the blue group furthest to the right" and to provide reasoning from a multiple choice list and grade their certainty. The experiment set up is shown in @fig-right. The participants answers were then compared to a ground truth where the correct plot had a right shifted mean. By comparing the results to a ground truth statistic and marking participants as "wrong" or "right", the error from the participants that had an alternative interpretation to the concept of "furthest right" was conflated with the error from a the visualisation choice.

While this is a focus on uncertainty visualisation, these issues in uncertainty communication need to be acknowledged because they bleed into the uncertainty visualisation literature. Uncertainty visualisation are compared on their ability to communicate explicit risk, elicit trust, and prevent risk or ambiguity avoidance, continuing these issues in uncertainty communication at large.

Ambiguity-aversion, uncertainties risk-avoidance counterpart, has a similar problem. Ambiguity-aversion is the tendency of people to prefer to take on certain risks rather than unknown risks. It is essentially risk-aversion applied to risk. For this reason discussions of ambiguity-aversion have the same issues as the conversations about risk-aversion, where they do not consider ambiguity-aversion to be an estimate of the utility of decreased certainty. Assuming that participants should act as rational agents implies they should do as rational agents do and *ignore* the uncertainty information.

The idea that risk and ambiguity aversion reveal the utility of decreased variance seems to seldom be considered in the literature, therefore it should be mentioned here, otherwise this issue might continue. However, considering risk aversion is at its core, a trade off between bias and variance, it is bizzare to assume it to be a mistake when statisticians perform these trade offs all the time. If risk-aversion is truly "irrational" behaviour, every statistical textbook that discusses the use of a biased but consistent estimator in a finite sample case (a common example would be specific cases of the maximum likelihood estimator) should be tossed out. Discussing risk and ambiguity aversion as though it is some diversion from a perfectly rational, mathematical choice, implies there is *no value to a decrease in variance which is not even true within mathematics*. 
There are some elements of risk communication that seem to translate to uncertaitny communication more broadly. Using words to communicate risk is discouraged because people can misinterpret the actual risk involved, so when possible numerical estimates of risk should be communicated [@Spiegelhalter2017] the same rules apply to uncertainty (*Cite gap __*). 
People performing better and preferring with numerical estimates of risk translates to uncertainty communication more broadly . Risk communication also needs to have clear objectives, use plain language, limit information to only what is necessary, and segment the audience to allow for differences in interest and knowledge [@Spiegelhalter2017]. These general concepts of communication also extend to communication of uncertainty. On the other hand, it is not sensible to assume that probability specific concepts such as framing, intuitiveness of probabilities, or reference classes are still relevant when we consider uncertainty communication as signal supression, not as a synonym of risk or probability communication. In this same vein, most uncertainty communication papers very often discuss risk-aversion as something that needs to be considered despite it being unclear exactly how risk-aversion is related to uncertainty communication *(citations for the papers that do this)*.

>>>>

---
title: "Writing Trash Bin"
author: "Harriet Mason"
format: html
---
# Introduction
- @Spiegelhalter2017 noted we "cannot assess the quality of risk communication unless the objectives are clear".
- @Fischhoff2014  takes this approach by considerting how decisions with different levels of flexibility should have different uncertainty visualisations. If the user has no choice they should simply be made aware of the risk, if they have several options, the user needs to see the uncertainty in each, if they need to come up with the options themselves, they need to see how things work and how the system fails.

### Motivation for uncert vis
- @Suh2023 found that the main issue with deployment of AI was communication between data science and subject matter experts, since subject matter experts didn't want technical presentations that didn't explain exactly when and how a model would fail.
- (maybe include some of the noise vs signal stuff)
- The use of uncertainty in high dimensional environments is especially important in energy data. Large models that incorporate spatial-temporal data from many sources and systems are used to predict energy uses in the short and long term. Understanding how to improve and make better decisions in these models is imperative in both the daily operation of the energy sector as well as in the transition from fossil fuels to clean energy. The energy sector needs better heuristics to make energy supply analysis less costly to conduct [@Stenclik2021], therefore it is an incredibly relevant application of uncertainty visualisation techniques.

# Definition Notes
- Failing to communicate the nature of your uncertainty can result in underestimation or overestimation of failure probabilities [@Kiureghian2009]

## General
- @uncertchap2022 People do not over react when an ensemble member barely misses their town and People overreact more when 1/9 enselbles appear to be hitting their town than when 1/33 ensembles.
- @uncertchap2022 Correll and Gleicher [9] found that violin plots (mapping probability to area or width at a given y position) and gradient plots (mapping probability to opacity) lead to more intuitive assessments of value likelihood and “ability to surprise,” which are more closely aligned with statistical defini- tions of uncertainty." wtf is the statistical definition of uncertainty
- @Thomson2005 titled their paper "A typology for visualizing uncertainty" however it is a taxonomy for the specific types of uncertainty that we may want to visualise and a way to calculate that uncertainty. While this paper is more about mathematics than visualisation, it is clearly written as a precursor to visualisation.
- @Griethe2006 commented that uncertainty appears at every step in the analyis pipeline, and commented on some the mathematical definitions that seek to quantify uncertianty, but did not explain how these mathematical definitions can be combined across levels. Unfortunately this quantification is needed since she mentioned that in order to visualise uncertainty, it needs to be measurable.
- mentions uncertainty can be variance, precision, accuracy, reliability, or related concepts @Hullman2016
- @uncertchap2022 "can enter the analysis pipeline during the mea- surement, modeling, and forecasting phases"

## Redefined uncertainty (cite gap 3)

#### Uncertainty Taxonomies (Cite Gap 7)

- @Gustafson2019 organises uncertainty into the four groups: Deficient: Lack of knowledge; Technical: modelling approximations and measurement error; Scientific Unknown unknowns; Consensus: Disagreement among parties
- @uncertchap2022 mention aleatric, epistemic, ontological divide and also differentiates risk with uncertainty
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016

@@

- @Griethe2006 clearly sees uncertainty as additional information or data, and therefore sees the uncertainty visualisation problem as a specific case of the high dimensional visualisation issues. 
- It is important to keep in mind that just because information is in a graphic, that does not mean it will be "seen". The phenomena of inattentional blindness shows that there is no perception without attention and it is powerful enough that participants can fail to be aware of random objects appearing on a screen or a gorilla walking through a basketball game [@simons1999gorillas; @mack2003inattentional]. Including uncertainty is worth very little if no attention is left to see it. 
- Glyph maps can be used to depict the multi-dimensional information in spatial-temporal data by mapping line plots to map locations, but we still need to decide what information is important. The map can either present trends in the global or local variance depending on whether or not the line plot is scaled globally or locally [@Wickham2012], however smaller details are almost impossible to convey. No matter how much information we try and put in a plot, there will always be only a handful of key takeaways.
- Elements from a single distribution should be displayed using a single plot, since displaying the features of one distribution across multiple plots makes the information hard to combine and results in some details (such as the estimate error) being completely ignored [@moritz2017trust; Correll2018].

# Noise vs Signal
- Upon further interviews @Hullman2020a found that authors believed uncertainty would overwhelm the audience and make their data seem questionable because decision makers are unable to understand uncertainty.
- @uncertchap2022 "Effectively communicating uncertainty is necessary for establishing scientific transparency"
- @Griethe2006 defined uncertainty as "the degree to which the lack of knowledge about the amount of error is responsible for hesitancy in accepting results and observations without caution". 
- confidence intervals provided in text form only are less likely to be misinterpreted than graphics [@Savelli2013], text is insufficient to express more complicated aspects of a distribution, such as mass.
  - This kind of harks back to the "if our main concern is accuracy, we should show the values in a table" issue.
- A survey conducted by @Hullman2020a found that majority of visualisation authors agreed that expressing uncertainty is important and should be done more often than it currently is, some even agreed that failing to do so is tantamount to fraud. Despite this, only a quarter of respondents included uncertainty in 50% or more of their visualisations [@Hullman2020a].
- Some economists suggest that visualisation authors are responding to incentives that make it tempting to avoid visualising uncertainty, even if those incentives are based more in perception than reality [@Manski2020].
- The study by @Hullman2020a found that the most common reasons authors don't visualise uncertainty despite knowing it's moral importance are: not wanting to overwhelm the audience; an inability to calculate the uncertainty; a lack of access to the uncertainty information; and not wanting to make their data seem questionable [@Hullman2020a].
- At least one interviewee from the study by @Hullman2020a claimed that expertise implies that the signal being conveyed is significant, but also said they would omit uncertainty if it obfuscated the message they were trying to convey.  Other authors who were capable of calculating and and representing uncertainty well did not do it, and were unable to provide a self-satisfying reason why [@Hullman2020a]. These conflicting motivations are acknowledged in the paper itself where @Hullman2020a says:

@@

### Value Extraction (Cite Gap 15)
- This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016].
- @uncertchap2022 Attribute substitution: people will substitute a hard mental calculation for an easy one (attribute substitution heuristic). Name "deterministic constructual error" when people try to substitute uncertainty information for deterministic information (e.g. error bars for high and low temperature even when the visualisation was explained - from icing study. This makes sense because temperature is GIVEN as high and lows so it is a natural interpretation). I think this result is in line with the idea that people give the best estimate based on what the visualisation suggests. i.e. people might have assumed you meant "what is the lowest possible temperature". There is a level of "asking a deterministic question for a probabilistic question". 
- @uncertchap2022 Quantile dot plot created so people dont have to visually calculate the area under the curve (so authors are aware of the mental challenge of some of these things) Decrease variance with respect to density plots (it makes sense because you can directly count and quantify raher than guesstimate) (both come from the same source, when-ish is my bus)
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016 
- @Correll2018 “place your 5 ships on the safest locations on the board.” (minimise danger) It is a convoluted value extraction task, where participants basicaly have to look at the palette and see that if it is in the lower uncertainty, and lower error it is better, but higher uncertainty is better
- @Ibrekk1987 expected participants to calculate the mean using a pie chart that had numerical bins combined (e.g. segment one was 0-2 inches of snow, and segment two was 2-4 inches of snow) so it could not be done by calculation. The authors also expected participants to use a CDF to calculate the mean, which needs to be converted *back* into a PDF to do so mathematically, so it was unclear how participants were supposed to do this calculation visually unless they were to just make a blind guess. The authors interpreted this as the participants misinterpreting how the CDF works, not seeming to realise they gave them an impossible task.
- @Sanyal2009 users performed significantly well using Glyph-size when the task was to search for locations of least uncertainty. However, both Glyph-color and Surface-color performed better than Glyph- size when the task was to search for locations of high uncertainty. We did not expect to find a significant difference between the two search tasks since both the tasks were designed to find extremes in the data
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014

### Decision making stuff (Cite Gap 16)
- Providing a categorical decision alone is somewhat useless [@Joslyn2012], and visualising a single estimate is akin to providing a decision or expressing no uncertainty at all. 
- This reality cannot be avoided by providing secondary or non-specific information such as explaining calculations [@Joslyn2012], explaining the advantages of a recommendation [@Joslyn2012], or expressing uncertainty in vague terms [@Erev_1990; @Olson_1997], all of which are undesirable for decision makers and lead to measurably worse decisions [@Joslyn2012; @Erev_1990; @Olson_1997].
- There is some research that suggests laypeople cannot understand complicated concepts in statistical thinking (such as trick questions on hypothesis tests or the difference between Frequentist and Bayesian thinking) [@Hoekstra2014; @Bella2005] but there is a large amount of research suggesting that presenting uncertainty information improves decision making, both experimentally [@Joslyn2012; @Savelli2013; @Kay2016; @Fernandes2018] and in practice [@Al-Kassab2014]. As a matter of fact, doing what many authors currently do (providing only a deterministic outcome with no uncertainty) causes decision makers to be *less* decisive and have completely unbounded expectations on an outcome [@Savelli2013].

- @Boone2018 organised responses from participants by the approach they used
- @Hullman2016 suggested authors look for signs of heuristics in responses and ask subjects to describe their strategy in their response. 
- @Hullman2016 What was particularly interesting was people's tendancy to ignore uncertainty in a plot all together, and use a pattern in the signal (such as distance between means) to answer questions that require uncertainty information. 
- @Kim2019 believed an understanding of participants prior beliefs could assist in interpreting uncertainty visualisation results that tend to be incredibly noisy.
- @Daradkeh2015 found that people tended to to make decisions on a small number of salient pieces of information rather than on systematic evaluations of the costs and benefits of choices.
- @Daradkeh2015 found that people tended to ignore or give little weight probability information and instead focused on the severity of undersierable outcomes.
- believed that offering enough decisions with varying rewards you can observe the true subjective probability, however whether or not probability is the right representation of subjective uncertainty is a topic of debate @Hullman2016
- she also discusses graphical interface (users have to draw the probability distribution) and framing probabilities as natural frequencies are better choices because they reduce noise in responses [@Hullman2016]
- Consider incorporating utility functions. Include "probability-coherence" checks [@Hullman2016]
- evidence that people do not follow the axioms of probability when conveying subjective probabilities. How do you deal with this as an analyst (e.g. if the probabilities do not sum to 1 @Hullman2016
- people will discount information they perceive as discountable if they are able. Including mean estimates on depictions of mass can cause people to discount the uncertainty information and use the difference between means as a proxy for the probability distribution [@Kale2021].
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016

@@

### Trust stuff (Cite Gap 18)
- @uncertchap2022  mentions "other studies have found that people trust icon arrays more than other common visualization techniques" without commenting on whether or not that trust is good?
- A 6-month survey of anti-mask groups on Facebook during the COVID-19 pandemic showed that the anti-maskers thought carefully about their grammar of graphics and made persuasive visualisations using the same data as pro-mask groups. They did this by exploiting information ignored by the pro-maskers [@Lee2021].
- displaying numerical estimates of uncertainty information has shown to lead to greater trust in predictions [@Joslyn2012; @VanderBles2020]. While @Han2009 found people have more worry when presented with uncertainty regarding health outcomes, this worry is not a bad thing if the concern is warranted given the ambiguous situation.

# Communication considerations
## Uncertainty
- Expressing uncertainty verbally decreases the perceived reliability and trustworthiness of the source [@VanderBles2020].
- @Gustafson2019 found that the framing of our uncertainty, (i.e. if the source of uncertainty is from a lack of knowledge, approximations, unknown unknowns, or disagreement among parties) was found to not have a detrimental effect on the belief in the estimates, perceived credibility, or behavioural intentions of the decision makers.
- Risk communication also needs to have clear objectives, use plain language, limit information to only what is necessary, and segment the audience to allow for differences in interest and knowledge [@Spiegelhalter2017].
- @Spiegelhalter2017 it is better to understand your audience and iterate towards a final design, rather than only consider one option.
- Some markers of probability that have common uses, probabilities such as 0 (impossible), 0.5 (a fair coin flip), 1 (certain) are easy for people to have an internal sense of [@Hullman2016]
- It is incredibly hard to communicate small risks, because people cannot differentiate between 1/100 an 1/1000, and communicating low probability but high impact risks are particularly challenging [@Spiegelhalter2017]. This issue can be somewhat alleviated by communicating the likelihood of an event in relation to another event on a similar scale (such as being struck by lightning) [@Spiegelhalter2017]. 
- The framing of how we communicate risks matters due to the affect heuristic, so there will be a difference in peoples behaviours if, for example, you tell people they have a 90% chance of survival vs a 10% chance of death, and when expressing a frequency, a larger numerator communicates a larger risk [@Spiegelhalter2017]. 
- The reference class of a particular risk should be explicitly stated to avoid confusion, for example if we are communicating the chance of rain, the time period over which we are expecting this chance of rain (i.e. 50% chance today or this week) needs to be specified [@Spiegelhalter2017]. 
- industrial accidents have a "tollerable" risk chart based on risk and number of fatalities @Spiegelhalter2017
- People believe in the law of small numbers but dont believe in the lar of large numbers @Hullman2016
- experiencing an event makes your percieved probability higher than if you see a description of it @Hullman2016
- Frequency thing means that more effective communication to participants in a study means more effective science and we have a roll on effect @Hullman2016
- Probability judgements are less relative than other psychophysical judgements, 0%, 50% and 100% all have well understood meanings. Subjective probability distributions are spontaneous, and subjects are constructing a distribution upon being asked rather than articulating a fully formed mental representation of their beliefs. Therefore it is subject to heuristics. consider anchoring a scale with farmiliar probabilities . not sure if being asked for the entire distribution helps, e.g. the graphical one reduces noise but asking people for an interval estimation leads to overprecision and overconfidence @Hullman2016
- The big data paradox shrinks confidence intervals but magnifies bias [@Bradley2021] because the confidence intervals shrink due to the sample size but the imprecise target group creates bias.

## Visualisation
- Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012].
- There is a reasonable amount of evidence that cumulative displays or discrete displays (such as a quantile dot-plots or histograms) are the best ways to express mass for decision making and probability estimates [@Fernandes2018; @Hofmann2012; Kay2016; @Hullman2018; @kale2019decision]. 
- Visualisation allow for interactive graphics that provide a more in depth understanding of probability [@Potter2009; @Ancker2009] and 
- Infographics that make uncertainty more accessible for people with poor numeracy skills [@Ancker2009]. 
- Any confusion caused by expressing uncertainty as a visualisation could also be due to a lack of exposure, since @Kay2016 found people repeatedly exposed to the same uncertainty visualisations quickly get better at making judgements.
- Even something as simple as sketching a distribution before recalling statistics or making predictions can greatly increase the accuracy of those measures [@Hullman2018; @Goldstein2014].
- @uncertchap2022 Icon array and quantile dot plots both support area and frequency based judgements.
- @uncertchap2022 frequency framing: (e.g. 1 out of 10 rather than 10%) better (plot: icon array, quantile dot plot). 
- @Spiegelhalter2017 even if most audience members can read a plot, they may not read it in the same way. People with high numeracy skills count pictographs, but people with low numeracy assess area, and seemingly unrelated elements may not equally influence all these methods of reading, for example scattering in a pictograph makes counting particularly more difficult. *also add that perceptual task audience thing*
- @Spiegelhalter2017 common sentiments such as avoiding chart junk, considering a table in place of a visualsiaiton, use of narrative labels, managing emotion, less is more approach, etc, are the main concepts expressed. 
- @Spiegelhalter2017 suggests that the graphic should allow "part-to-whole" comparison on an appropriate scale, which seems to be an extension of the framing concept discussed when speaking about uncertainty communication more broadly.
- @Hullman2016 discussed the role of heuristics in the analysis of results
- @Kinkeldey2014 stress the importance of selecting the correct audience, as numerical and visual literacy can differ vaslty between subpopulations so the results from one populations may not reflect the results from another. 
- we consider a summary table as a visualisation, illuminate graphics with words and numbers, design graphics to allow a part-to-whole comparison on an appropriate scale, use images and metaphors to gain/retain attention but don't arouse undue emotion, assume low numeracy of a general public audience and adopt a less-is-more approach by reducing the need for inference and making clear and explicit comparison, be cautious about interactivity and animations (may introduce unnecessary complexity), avoid chart junk (like 3d bar charts) @Spiegelhalter2017
- @Maceachren2012 found __????_ to be the worst aesthetics to map to uncertainty to as they don't have an intuitive interpretation.
- Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. 
- No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012]. 
- This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016].
- coloured maps for earthquake risk are easily interpretable @Spiegelhalter2017
